From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-17 Pi+/paper/finite.tex

However, more conveniently, in HoTT, we can instead use groupoids to talk about groups. A group can be identified with a
1-object groupoid, using a technique called delooping. The delooping of a group $G$ is a groupoid $\B{G}$ given by a
unique object $\pt$ with self-loops that are 1-paths $\pt \id_{\B{G}} \pt$ corresponding to the elements of $G$. Note
that the group operations are given by operations on the identity type automatically, with $\refl_{\pt}$ for the neutral
element, path composition for the group multiplication, and path inverse for the group inverse. These satisfy the group
laws as well, up to the identity type, using the groupoid coherence laws. Moreover, for 1-groups which are supposed to
be sets, these 2-paths should be propositions, so we have to restrict $\B{G}$ to be a 1-groupoid. Hence, a group is
simply given by a pointed, connected 1-type~\cite*{buchholtzHigherGroupsHomotopy2018,symmetryBook2021}.

For example, given a pointed type $(A:\UU, a:A)$, the automorphism group structure at $a$ is given by $a \id_{A} a$. Of
course, for 1-groups we will require that $a \id_{A} a$ is an $\hSet$, which is enforced by having $A$ be a groupoid. In
our running example for the permutation group on finite sets, we have that
$\Aut[\Fin[n]] \defeq (\Fin[n] \eqv \Fin[n]) \eqv (\Fin[n] \id_{\UU} \Fin[n])$, and the corresponding pointed, connected
groupoid is given by the classifying type $\UFin[n] \defeq \BAut[\Fin[n]]$, with
$\loopspace[\BAut[\Fin[n]],F_{n}] \eqv \Aut[\Fin[n]]$ as the corresponding automorphism group. \vc{revisit}

>>>>

However, more conveniently, in HoTT, we can instead use groupoids to talk about groups. A group can be identified with a
1-object groupoid, using a technique called delooping. The delooping of a group $G$ is a groupoid $\B{G}$ given by a
unique object $\pt$ with self-loops that are 1-paths $\pt \id_{\B{G}} \pt$ corresponding to the elements of $G$. Note
that the group operations are automatically given by operations on the identity type, with $\refl_{\pt}$ for the neutral
element, path composition for the group multiplication, and path inverse for the group inverse. These satisfy the group
laws as well, up to the identity type, using the groupoid coherence laws. Moreover, for 1-groups which are supposed to
be sets, these 2-paths should be propositions, so we have to restrict $\B{G}$ to be a 1-groupoid. Hence, a group is
simply given by a pointed, connected 1-type~\cite*{buchholtzHigherGroupsHomotopy2018,symmetryBook2021}.

For example, given a pointed type $(A:\UU, a:A)$, the automorphism group structure at $a$ is given by $a \id_{A} a$. Of
course, for 1-groups we will require that $a \id_{A} a$ is an $\hSet$, which is enforced by having $A$ be a groupoid. In
our running example for the permutation group on finite sets, we have that $\Fin[n]$ is an $\hSet$, and hence,
$\UFin[n] \defeq \BAut[\Fin[n]]$ is a pointed, connected 1-type, whose loopspace $\loopspace[\BAut[\Fin[n]],F_{n}]$ is
equivalent to $\Aut[\Fin[n]] \defeq (\Fin[n] \eqv \Fin[n]) \eqv (\Fin[n] \id_{\UU} \Fin[n])$, which has the
corresponding automorphism group structure.
